---
title: "Generalized Linear Mixed Models for Moderator Analyses"
date: "2025-07-20"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    toc_depth: 4
    number_sections: true
    code_folding: hide
  word_document:
    toc: true
    toc_depth: 4
  pdf_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = FALSE, cache = FALSE)

# Set the root directory for the entire document (recommended way)
knitr::opts_knit$set(root.dir = dirname(rstudioapi::getActiveDocumentContext()$path))

```

# Preparation

```{r load-packages}
library(tidyverse)
library(lme4)
library(broom.mixed)
library(forcats)
library(gt)
library(knitr)
```

## Load Data

```{r load-data}
# load data
df_moderation <- read.csv2("../pre-processing/output/data_for_moderation_analyses.csv", dec = ".")

# calculate Sample_Size_100
df_moderation <- df_moderation %>%
  mutate(Sample_Size_100 = Sample_Size / 100)

# Display the first rows to check the data
head(df_moderation)
```


## Descriptive Statistics

We provide basic descriptive statistics of the included studies, without excluding any data due to missingness.

```{r descriptives}

# Calculate descriptives
total_number_of_samples <- nrow(df_moderation)
total_sample_size <- sum(df_moderation$Sample_Size, na.rm = TRUE)

# Create a summary table
summary_table <- data.frame(
  Description = c("Number of studies / cohorts", "Total sample size (sum of available Sample_Size)"),
  Value = format(c(total_number_of_samples, total_sample_size), big.mark = ",")
)

# Display as a table
knitr::kable(summary_table, caption = "Descriptive Statistics of the Dataset")

```
# Moderator Analyses

Here we fit separate binomial GLMMs (random intercept for `Study`) for each moderator. For continuous moderators, we report the estimated prevalence at the reference (moderator = 0) and the marginal delta (approximate change in percentage points per 1‑unit increase). For categorical moderators, we report the estimated prevalence for the reference level and contrasted level(s).

## Moderators setup
```{r mods-setup-functions}
# distinguish between continuous and categorical variables
cont_vars <- c(
  "Percentage_women",
  "Mean_age",
  "Percentage_minority",
  "Percentage_partner",
  "High_education",
  "TP_assessments",
  "N_trajectories",
  "Trauma_TP1",
  "TP1_TPX",
  "Trauma_TPX",
  "Entropy",
  "Grolts",
  "Sample_Size_100" # unit = 100 participants
)

cat_vars <- list(
  list(var="Diagnostic_DSM",     keep=c("4","5"),                 ref="4"),
  list(var="Occupational_trauma",keep=c("No","Yes"),              ref="No"),
  list(var="Developmental_age",  keep=c("Adult","Youth"),         ref="Adult"),
  list(var="Location",           keep=NULL,                       ref=NULL),
  list(var="Location_US",        keep=NULL,                       ref="Non-US"),
  list(var="Trajectory_analysis",keep=NULL,                       ref=NULL),
  list(var="Scale_moderator",    keep=c("Interview","Self"),      ref="Interview"),
  list(var="Trauma_exposure",    keep=c("Inter","Non"),           ref="Inter"),
  list(var="Trauma_type",        keep=c("Combat","Natural","Injury"), ref="Combat"),
  list(var="Military",           keep=NULL,                       ref="No"),
  list(var="Discrete",           keep=c("No","Yes"),              ref="No"),
  list(var="Health_First",       keep=c("Yes","No"),              ref="Yes")
)

# Consistent optimizer for all fits
ctrl <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))

# Utility to add Success/Failure columns for a given trajectory
add_binom_cols <- function(data, success_col = "Low_n", denom_col = "Sample_Size") {
  stopifnot(success_col %in% names(data), denom_col %in% names(data))
  data %>%
    mutate(
      Success = .data[[success_col]],
      Failure = .data[[denom_col]] - Success
    )
}

#fit model for one predictor and return the key stats
fit_cont <- function(data, var, success_col = "Low_n", denom_col = "Sample_Size") {
  d <- data %>%
    filter(!is.na(.data[[var]])) %>%
    add_binom_cols(success_col = success_col, denom_col = denom_col)

  fml <- as.formula(paste0("cbind(Success, Failure) ~ ", var, " + (1 | Study)"))
  m <- glmer(fml, data = d, family = binomial, control = ctrl)

  coefs <- summary(m)$coefficients
  b0 <- unname(fixef(m)["(Intercept)"])
  b1 <- unname(fixef(m)[var])
  p0 <- plogis(b0)

  p_col <- intersect(colnames(coefs), c("Pr(>|z|)", "Pr(>|t|)", "Pr(>|Chisq|)"))
  pval  <- if (length(p_col)) unname(coefs[var, p_col[1]]) else NA_real_

  tibble(
    trajectory = sub("_n$", "", success_col),
    predictor  = var,
    Samples    = nobs(m),
    prevalence_at_0_pct        = 100 * p0,
    slope                      = b1,
    marginal_delta_pp_per_unit = b1 * p0 * (1 - p0) * 100,
    p_value                    = pval
  )
}

# Fit one categorical moderator and return prevalence by level + Δ vs ref (pp) + p-values
fit_cat <- function(data, var, keep = NULL, ref = NULL,
                    success_col = "Low_n", denom_col = "Sample_Size") {
  stopifnot(var %in% names(data))

  d <- data %>%
    filter(!is.na(.data[[var]])) %>%
    { if (!is.null(keep)) filter(., .data[[var]] %in% keep) else . } %>%
    mutate(
      !!var := factor(.data[[var]]),
      !!var := if (!is.null(ref) && ref %in% levels(.data[[var]]))
                 fct_relevel(.data[[var]], ref)
               else fct_drop(.data[[var]])
    ) %>%
    add_binom_cols(success_col = success_col, denom_col = denom_col)

  if (nlevels(d[[var]]) < 2) return(tibble())  # nothing to compare

  fml <- as.formula(paste0("cbind(Success, Failure) ~ ", var, " + (1 | Study)"))
  m <- glmer(fml, data = d, family = binomial, control = ctrl)

  fe <- fixef(m)
  b0 <- unname(fe["(Intercept)"])
  lvls <- levels(d[[var]])

  # baseline = first level
  add <- sapply(lvls, function(L) {
    nm <- paste0(var, L)                 # e.g., "Trauma_typeNatural"
    if (nm %in% names(fe)) fe[[nm]] else 0
  })

  s <- summary(m)$coefficients
  pcol <- intersect(colnames(s), c("Pr(>|z|)", "Pr(>|t|)", "Pr(>|Chisq|)"))
  pvals <- sapply(lvls, function(L) {
    nm <- paste0(var, L)
    if (nm %in% rownames(s)) unname(s[nm, pcol[1]]) else NA_real_  # ref has NA
  })

  p <- plogis(b0 + add)
  p_ref <- p[1]

  tibble(
    trajectory      = sub("_n$", "", success_col),
    moderator       = var,
    n               = nobs(m),               # Samples (k)
    ref_level       = lvls[1],
    level           = lvls,
    prevalence_pct  = 100 * p,
    diff_vs_ref_pp  = 100 * (p - p_ref),
    p_value         = pvals
  )
}
```
## Tables Setup

```{r mods-setup-tables}
# helpers for p formatting
fmt_p <- function(p) ifelse(is.na(p), "", ifelse(p < .001, "< .001", sprintf("%.3f", p)))
sig   <- function(p) dplyr::case_when(is.na(p) ~ "", p < .001 ~ "***", p < .01 ~ "**", p < .05 ~ "*", TRUE ~ "")

# Moderator labels for tables
mod_labels <- c(
  Percentage_women    = "Percentage of women",
  Mean_age            = "Mean age",
  Percentage_minority = "Minorities (%)",
  Percentage_partner  = "Having a partner (%)",
  High_education      = "Education (%)",
  TP_assessments      = "Time point PTSD assessments",
  N_trajectories      = "Number of trajectories",
  Trauma_TP1          = "Time between trauma and first assessment (months)",
  TP1_TPX             = "Time span of the study (months)",
  Trauma_TPX          = "Time between trauma and last assessment (months)",
  Entropy             = "Entropy",
  Grolts              = "Quality score",
  Sample_Size_100     = "Sample size / 100",
  Diagnostic_DSM      = "Diagnostic Criteria",
  Occupational_trauma = "Occupational trauma",
  Developmental_age   = "Developmental age",
  Location            = "Country income",
  Location_US         = "US vs Non-US",
  Trajectory_analysis = "LGMM vs LCGA",
  Scale_moderator     = "PTSD symptom assessment",
  Trauma_exposure     = "Trauma exposure",
  Trauma_type         = "Trauma nature",
  Military            = "Civilian vs Military",
  Discrete            = "Discrete vs non-discrete trauma",
  Health_First        = "HP/First responders vs other"
)

# continuous variable labels for tables
mod_levels <- unname(mod_labels)
delta_col  <- "Marginal Δ"  # use literal Δ; save file as UTF-8

# categorical moderator level labels for tables
level_labels <- list(
  Diagnostic_DSM      = c(`4` = "DSM-IV", `5` = "DSM-5"),
  Occupational_trauma = c(No = "No", Yes = "Yes"),
  Developmental_age   = c(Adult = "Adult", Youth = "Youth"),
  Location            = c(
    "High income countries" = "High income countries",
    "High-income countries" = "High income countries",
    "High"                  = "High income countries",
    "HIC"                   = "High income countries",
    "Mid-, low- income"     = "Mid-, low- income",
    "Mid/low income"        = "Mid-, low- income",
    "LMIC"                  = "Mid-, low- income"
  ),
  Location_US         = c(`Non-US` = "Non-US", US = "US"),

  Trajectory_analysis = c(LCGA = "LCGA", LGMM = "LGMM"),
  Scale_moderator     = c(Interview = "Interview", Self = "Self-reported scale"),
  Trauma_exposure     = c(Inter = "Interpersonal trauma", Non = "Non-interpersonal trauma"),
  Trauma_type         = c(Combat = "Combat", Injury = "Injury", Natural = "Natural disaster"),
  Military            = c(No = "Civilian", Yes = "Military"),
  Discrete            = c(No = "Non-discrete", Yes = "Discrete"),
  Health_First        = c(Yes = "HP/First responder", No = "Other")
)

# Build a lookup tibble we can join onto both ref and comparison
lvl_map <- purrr::imap_dfr(level_labels, ~tibble::tibble(
  moderator = .y,
  orig      = names(.x),
  pretty    = unname(.x)
))
```

## 🚀 Low Symptom Trajectory

```{r low-moderators-setup}

df_Low <- df_moderation %>%
  filter(!is.na(Low_percentage))

# Ensure Low_percentage is numeric
df_Low$Low_percentage <- as.numeric(df_Low$Low_percentage)

# Calculate Low_n as (percentage / 100) * sample size
df_Low$Low_n <-
  round((df_low$Low_percentage / 100) * df_Low$Sample_Size)


```
### Descriptives

We provide additional descriptives based on the Low trajectory variable, using the full dataset.

```{r Low-descriptives}

# Total number of individuals classified as Low (ignoring missingness in other variables)
total_Low_n <- sum(df_Low$Low_n, na.rm = TRUE)

# Number of unique studies (assumes a column Study exists)
unique_studies <- length(unique(df_Low$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in Low trajectory", "Number of unique studies"),
  Value = format(c(total_Low_n, unique_studies), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for Low Trajectory Analysis")
```
### Continuous moderators

```{r Low-mods-cont}
fit_cont_Low <- \(data, var) fit_cont(data, var, success_col = "Low_n", denom_col = "Sample_Size")

fits_cont_Low <- map_dfr(cont_vars, ~fit_cont_Low(df_Low, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# table rendering
tbl <- fits_cont_Low %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

tbl_utf8 %>%
  gt() %>%
  tab_header(title = "Low Symptoms Trajectory - Continuous Moderators") %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r Low-mods-cat}
fit_cat_Low <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "Low_n", denom_col = "Sample_Size")

fits_cat_Low <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_Low, c(list(data = df_Low), spec))
)

# table
## set reference level
ref_prev_low <- fits_cat_Low %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

# setup table
Low_cat_table <- fits_cat_Low %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for LOW
  dplyr::left_join(ref_prev_low, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

# Render
Low_cat_table %>%
  gt() %>%
  tab_header(title = "Low Symptoms Trajectory — Categorical Moderators") %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```

## 🌿 Decreasing Symptom Trajectory

```{r Decreasing-moderators-setup}

df_Decreasing <- df_moderation %>%
  filter(!is.na(Decreasing_percentage))

# Ensure Decreasing_percentage is numeric
df_Decreasing$Decreasing_percentage <- as.numeric(df_Decreasing$Decreasing_percentage)

# Calculate Decreasing_n as (percentage / 100) * sample size
df_Decreasing$Decreasing_n <-
  round((df_Decreasing$Decreasing_percentage / 100) * df_Decreasing$Sample_Size)
```

### Descriptives

We provide additional descriptives based on the Decreasing trajectory variable, using the full dataset.

```{r Decreasing-descriptives}
# Total number of individuals classified as Decreasing (ignoring missingness in other variables)
total_Decreasing_n <- sum(df_Decreasing$Decreasing_n, na.rm = TRUE)

# Number of unique studies (assumes a column Study exists)
unique_studies <- length(unique(df_Decreasing$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in Decreasing trajectory", "Number of unique studies"),
  Value = format(c(total_Decreasing_n, unique_studies), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for Decreasing Trajectory Analysis")
```
### Continuous moderators

```{r Decreasing-mods-cont}
fit_cont_Decreasing <- \(data, var) fit_cont(data, var, success_col = "Decreasing_n", denom_col = "Sample_Size")

fits_cont_Decreasing <- map_dfr(cont_vars, ~fit_cont_Decreasing(df_Decreasing, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# table 
# build table
tbl <- fits_cont_Decreasing %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

# render
tbl_utf8 %>%
  gt() %>%
  tab_header(title = "Decreasing Symptoms Trajectory - Continuous Moderators") %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r Decreasing-mods-cat}
fit_cat_Decreasing <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "Decreasing_n", denom_col = "Sample_Size")

fits_cat_Decreasing <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_Decreasing, c(list(data = df_Decreasing), spec))
)

# table
## build table 
ref_prev_Decreasing <- fits_cat_Decreasing %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

Decreasing_cat_table <- fits_cat_Decreasing %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for Decreasing
  dplyr::left_join(ref_prev_Decreasing, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

## Render
Decreasing_cat_table %>%
  gt() %>%
  tab_header(title = "Decreasing Symptoms Trajectory — Categorical Moderators") %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```

## ⚠️ Increasing Symptom Trajectory

```{r Increasing-moderators-setup}

df_Increasing <- df_moderation %>%
  filter(!is.na(Increasing_percentage))

# Ensure Increasing_percentage is numeric
df_Increasing$Increasing_percentage <- as.numeric(df_Increasing$Increasing_percentage)

# Calculate Increasing_n as (percentage / 100) * sample size
df_Increasing$Increasing_n <-
  round((df_Increasing$Increasing_percentage / 100) * df_Increasing$Sample_Size)

```
### Descriptives

We provide additional descriptives based on the Increasing trajectory variable, using the full dataset.

```{r Increasing-descriptives}

# Total number of individuals classified as Increasing (ignoring missingness in other variables)
total_Increasing_n <- sum(df_Increasing$Increasing_n, na.rm = TRUE)

# Number of unique studies (assumes a column Study exists)
unique_studies <- length(unique(df_Increasing$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in Increasing trajectory", "Number of unique studies"),
  Value = format(c(total_Increasing_n, unique_studies), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for Increasing Trajectory Analysis")
```
### Continuous moderators

```{r Increasing-mods-cont}
fit_cont_Increasing <- \(data, var) fit_cont(data, var, success_col = "Increasing_n", denom_col = "Sample_Size")

fits_cont_Increasing <- map_dfr(cont_vars, ~fit_cont_Increasing(df_Increasing, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# build table
tbl <- fits_cont_Increasing %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

tbl_utf8 %>%
  gt() %>%
  tab_header(title = "Increasing Symptoms Trajectory - Continuous Moderators") %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r Increasing-mods-cat}
fit_cat_Increasing <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "Increasing_n", denom_col = "Sample_Size")

fits_cat_Increasing <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_Increasing, c(list(data = df_Increasing), spec))
)

# build table
ref_prev_Increasing <- fits_cat_Increasing %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

Increasing_cat_table <- fits_cat_Increasing %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for Increasing
  dplyr::left_join(ref_prev_Increasing, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

# render
Increasing_cat_table %>%
  gt() %>%
  tab_header(title = "Increasing Symptoms Trajectory — Categorical Moderators") %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```

## 🩸 High Symptom Trajectory

```{r High-moderators-setup}

df_High <- df_moderation

df_High <- df_High %>%
  filter(!is.na(High_percentage))

# Calculate High_n as (percentage / 100) * sample size
df_High$High_n <-
  round((df_High$High_percentage / 100) * df_High$Sample_Size)

# Ensure High_percentage is numeric
df_High$High_percentage <- as.numeric(df_High$High_percentage)


```
### Descriptives

We provide additional descriptives based on the High trajectory variable, using the full dataset.

```{r High-descriptives}

# Total number of individuals classified as High (ignoring missingness in other variables)
total_High_n <- sum(df_High$High_n, na.rm = TRUE)

# Number of unique studies (assumes a column Study exists)
unique_studies <- length(unique(df_High$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in High trajectory", "Number of unique studies"),
  Value = format(c(total_High_n, unique_studies), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for High Trajectory Analysis")
```
### Continuous moderators

```{r High-mods-cont}
fit_cont_High <- \(data, var) fit_cont(data, var, success_col = "High_n", denom_col = "Sample_Size")

fits_cont_High <- map_dfr(cont_vars, ~fit_cont_High(df_High, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# build table
tbl <- fits_cont_High %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

# render
tbl_utf8 %>%
  gt() %>%
  tab_header(title = "High Symptoms Trajectory - Continuous Moderators") %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r High-mods-cat}
fit_cat_High <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "High_n", denom_col = "Sample_Size")

fits_cat_High <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_High, c(list(data = df_High), spec))
)

# build table
ref_prev_High <- fits_cat_High %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

High_cat_table <- fits_cat_High %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for High
  dplyr::left_join(ref_prev_High, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

# render
High_cat_table %>%
  gt() %>%
  tab_header(title = "High Symptoms Trajectory — Categorical Moderators") %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```

## 🌗 Moderate Symptom Trajectory

```{r Moderate-moderators-setup}

df_Moderate <- df_moderation


df_Moderate <- df_Moderate %>%
  filter(!is.na(Moderate_percentage))

# Calculate Moderate_n as (percentage / 100) * sample size
df_Moderate$Moderate_n <-
  round((df_Moderate$Moderate_percentage / 100) * df_Moderate$Sample_Size)

# Ensure Moderate_percentage is numeric
df_Moderate$Moderate_percentage <- as.numeric(df_Moderate$Moderate_percentage)


```
### Descriptives

We provide additional descriptives based on the Moderate trajectory variable, using the full dataset.

```{r Moderate-descriptives}

# Total number of individuals classified as Moderate (ignoring missingness in other variables)
total_Moderate_n <- sum(df_Moderate$Moderate_n, na.rm = TRUE)

# Number of unique studies (assumes a column Study exists)
unique_studies <- length(unique(df_Moderate$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in Moderate trajectory", "Number of unique studies"),
  Value = format(c(total_Moderate_n, unique_studies), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for Moderate Trajectory Analysis")
```
### Continuous moderators

```{r Moderate-mods-cont}
fit_cont_Moderate <- \(data, var) fit_cont(data, var, success_col = "Moderate_n", denom_col = "Sample_Size")

fits_cont_Moderate <- map_dfr(cont_vars, ~fit_cont_Moderate(df_Moderate, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# build table
tbl <- fits_cont_Moderate %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

tbl_utf8 %>%
  gt() %>%
  tab_header(title = "Moderate Symptoms Trajectory - Continuous Moderators") %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r Moderate-mods-cat}
fit_cat_Moderate <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "Moderate_n", denom_col = "Sample_Size")

fits_cat_Moderate <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_Moderate, c(list(data = df_Moderate), spec))
)

# build table
ref_prev_Moderate <- fits_cat_Moderate %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

Moderate_cat_table <- fits_cat_Moderate %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for Moderate
  dplyr::left_join(ref_prev_Moderate, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

# render
Moderate_cat_table %>%
  gt() %>%
  tab_header(title = "Moderate Symptoms Trajectory — Categorical Moderators") %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```
---











