---
title: "Generalized Linear Mixed Models for Moderator Analyses"
date: "2025-10-07"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    toc_depth: 4
    number_sections: true
    code_folding: hide
  word_document:
    toc: true
    toc_depth: 4
  pdf_document:
    toc: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = TRUE, message = FALSE, cache = FALSE)

# Set the root directory for the entire document (recommended way)
knitr::opts_knit$set(root.dir = dirname(rstudioapi::getActiveDocumentContext()$path))
```

# Preparation

```{r load-packages}
library(tidyverse)
library(lme4)
library(broom.mixed)
library(forcats)
library(gt)
library(knitr)
library(dplyr)
library(purrr)
library(stringr)
library(tidyr)
library(flextable)
library(officer)
```

## Load Data

```{r load-data}
# load data
df_moderation <- read.csv2("../pre-processing/output/data_for_moderation_analyses.csv", dec = ".")

# calculate Sample_Size_100
df_moderation <- df_moderation %>%
  mutate(Sample_Size_100 = Sample_Size / 100)

# Display the first rows to check the data
head(df_moderation)
```

## Descriptive Statistics

We provide basic descriptive statistics of the included studies, without excluding any data due to missingness.

```{r descriptives}
# Calculate descriptives
total_number_of_samples <- nrow(df_moderation)
total_sample_size <- sum(df_moderation$Sample_Size, na.rm = TRUE)

# Create a summary table
summary_table <- data.frame(
  Description = c("Number of samples / cohorts", "Total sample size (sum of available Sample_Size)"),
  Value = format(c(total_number_of_samples, total_sample_size), big.mark = ",")
)

# Display as a table
knitr::kable(summary_table, caption = "Descriptive Statistics of the Dataset")
```

# Moderator Analyses

Here we fit binomial GLMMs (random intercept for `Study`) separately for the continuous and categorical moderators. For continuous moderators, we report the estimated prevalence at the reference (moderator = 0) and the marginal delta (approximate change in percentage points per 1‑unit increase). For categorical moderators, we report the estimated prevalence for the reference level and contrasted level(s).

## Moderators setup
```{r mods-setup-functions}
# distinguish between continuous and categorical variables
cont_vars <- c(
  "Percentage_women",
  "Mean_age",
  "Percentage_minority",
  "Percentage_partner",
  "High_education",
  "TP_assessments",
  "N_trajectories",
  "Trauma_TP1",
  "TP1_TPX",
  "Trauma_TPX",
  "Entropy",
  "Grolts",
  "Sample_Size_100" # unit = 100 participants
)

cat_vars <- list(
  list(var="Diagnostic_DSM",     keep=c("4","5"),                 ref="4"),
  list(var="Occupational_trauma",keep=c("No","Yes"),              ref="No"),
  list(var="Developmental_age",  keep=c("Adult","Youth"),         ref="Adult"),
  list(var="Location",           keep=NULL,                       ref=NULL),
  list(var="Location_US",        keep=NULL,                       ref="Non-US"),
  list(var="Trajectory_analysis",keep=NULL,                       ref=NULL),
  list(var="Scale_moderator",    keep=c("Interview","Self"),      ref="Interview"),
  list(var="Trauma_exposure",    keep=c("Inter","Non"),           ref="Inter"),
  list(var="Trauma_type",        keep=c("Combat","Natural","Injury"), ref="Combat"),
  list(var="Military",           keep=NULL,                       ref="No"),
  list(var="Discrete",           keep=c("No","Yes"),              ref="No"),
  list(var="Health_First",       keep=c("Yes","No"),              ref="Yes")
)

# Consistent optimizer for all fits
ctrl <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5))

# Utility to add Success/Failure columns for a given trajectory
add_binom_cols <- function(data, success_col = "Low_n", denom_col = "Sample_Size") {
  stopifnot(success_col %in% names(data), denom_col %in% names(data))
  data %>%
    mutate(
      Success = .data[[success_col]],
      Failure = .data[[denom_col]] - Success
    )
}

#fit model for one predictor and return the key stats
fit_cont <- function(data, var, success_col = "Low_n", denom_col = "Sample_Size") {
  d <- data %>%
    filter(!is.na(.data[[var]])) %>%
    add_binom_cols(success_col = success_col, denom_col = denom_col)

  fml <- as.formula(paste0("cbind(Success, Failure) ~ ", var, " + (1 | Study)"))
  m <- glmer(fml, data = d, family = binomial, control = ctrl)

  coefs <- summary(m)$coefficients
  b0 <- unname(fixef(m)["(Intercept)"])
  b1 <- unname(fixef(m)[var])
  p0 <- plogis(b0)

  p_col <- intersect(colnames(coefs), c("Pr(>|z|)", "Pr(>|t|)", "Pr(>|Chisq|)"))
  pval  <- if (length(p_col)) unname(coefs[var, p_col[1]]) else NA_real_

  tibble(
    trajectory = sub("_n$", "", success_col),
    predictor  = var,
    Samples    = nobs(m),
    prevalence_at_0_pct        = 100 * p0,
    slope                      = b1,
    marginal_delta_pp_per_unit = b1 * p0 * (1 - p0) * 100,
    p_value                    = pval
  )
}

# Fit one categorical moderator and return prevalence by level + Δ vs ref (pp) + p-values
fit_cat <- function(data, var, keep = NULL, ref = NULL,
                    success_col = "Low_n", denom_col = "Sample_Size") {
  stopifnot(var %in% names(data))

  d <- data %>%
    filter(!is.na(.data[[var]])) %>%
    { if (!is.null(keep)) filter(., .data[[var]] %in% keep) else . } %>%
    mutate(
      !!var := factor(.data[[var]]),
      !!var := if (!is.null(ref) && ref %in% levels(.data[[var]]))
                 fct_relevel(.data[[var]], ref)
               else fct_drop(.data[[var]])
    ) %>%
    add_binom_cols(success_col = success_col, denom_col = denom_col)

  if (nlevels(d[[var]]) < 2) return(tibble())  # nothing to compare

  fml <- as.formula(paste0("cbind(Success, Failure) ~ ", var, " + (1 | Study)"))
  m <- glmer(fml, data = d, family = binomial, control = ctrl)

  fe <- fixef(m)
  b0 <- unname(fe["(Intercept)"])
  lvls <- levels(d[[var]])

  # baseline = first level
  add <- sapply(lvls, function(L) {
    nm <- paste0(var, L)                 # e.g., "Trauma_typeNatural"
    if (nm %in% names(fe)) fe[[nm]] else 0
  })

  s <- summary(m)$coefficients
  pcol <- intersect(colnames(s), c("Pr(>|z|)", "Pr(>|t|)", "Pr(>|Chisq|)"))
  pvals <- sapply(lvls, function(L) {
    nm <- paste0(var, L)
    if (nm %in% rownames(s)) unname(s[nm, pcol[1]]) else NA_real_  # ref has NA
  })

  p <- plogis(b0 + add)
  p_ref <- p[1]

  tibble(
    trajectory      = sub("_n$", "", success_col),
    moderator       = var,
    n               = nobs(m),               # Samples (k)
    ref_level       = lvls[1],
    level           = lvls,
    prevalence_pct  = 100 * p,
    diff_vs_ref_pp  = 100 * (p - p_ref),
    p_value         = pvals
  )
}
```

## Tables Setup

```{r mods-setup-tables}
# helpers for p formatting
fmt_p <- function(p) ifelse(is.na(p), "", ifelse(p < .001, "< .001", sprintf("%.3f", p)))
sig   <- function(p) dplyr::case_when(is.na(p) ~ "", p < .001 ~ "***", p < .01 ~ "**", p < .05 ~ "*", TRUE ~ "")

# Moderator labels for tables
mod_labels <- c(
  Percentage_women    = "Percentage of women",
  Mean_age            = "Mean age",
  Percentage_minority = "Minorities (%)",
  Percentage_partner  = "Having a partner (%)",
  High_education      = "Education (%)",
  TP_assessments      = "Time point PTSD assessments",
  N_trajectories      = "Number of trajectories",
  Trauma_TP1          = "Time between trauma and first assessment (months)",
  TP1_TPX             = "Time span of the study (months)",
  Trauma_TPX          = "Time between trauma and last assessment (months)",
  Entropy             = "Entropy",
  Grolts              = "Quality score",
  Sample_Size_100     = "Sample size / 100",
  Diagnostic_DSM      = "Diagnostic Criteria",
  Occupational_trauma = "Occupational trauma",
  Developmental_age   = "Developmental age",
  Location            = "Country income",
  Location_US         = "US vs Non-US",
  Trajectory_analysis = "LGMM vs LCGA",
  Scale_moderator     = "PTSD symptom assessment",
  Trauma_exposure     = "Trauma exposure",
  Trauma_type         = "Trauma nature",
  Military            = "Civilian vs Military",
  Discrete            = "Discrete vs non-discrete trauma",
  Health_First        = "HP/First responders vs other"
)

# continuous variable labels for tables
mod_levels <- unname(mod_labels)
delta_col  <- "Marginal Δ"  # use literal Δ; save file as UTF-8

# categorical moderator level labels for tables
level_labels <- list(
  Diagnostic_DSM      = c(`4` = "DSM-IV", `5` = "DSM-5"),
  Occupational_trauma = c(No = "No", Yes = "Yes"),
  Developmental_age   = c(Adult = "Adult", Youth = "Youth"),
  Location            = c(
    "High income countries" = "High income countries",
    "High-income countries" = "High income countries",
    "High"                  = "High income countries",
    "HIC"                   = "High income countries",
    "Mid-, low- income"     = "Mid-, low- income",
    "Mid/low income"        = "Mid-, low- income",
    "LMIC"                  = "Mid-, low- income"
  ),
  Location_US         = c(`Non-US` = "Non-US", US = "US"),

  Trajectory_analysis = c(LCGA = "LCGA", LGMM = "LGMM"),
  Scale_moderator     = c(Interview = "Interview", Self = "Self-reported scale"),
  Trauma_exposure     = c(Inter = "Interpersonal trauma", Non = "Non-interpersonal trauma"),
  Trauma_type         = c(Combat = "Combat", Injury = "Injury", Natural = "Natural disaster"),
  Military            = c(No = "Civilian", Yes = "Military"),
  Discrete            = c(No = "Non-discrete", Yes = "Discrete"),
  Health_First        = c(Yes = "HP/First responder", No = "Other")
)

# Build a lookup tibble we can join onto both ref and comparison
lvl_map <- purrr::imap_dfr(level_labels, ~tibble::tibble(
  moderator = .y,
  orig      = names(.x),
  pretty    = unname(.x)
))
```

## 🚀 Low Symptom Trajectory

```{r low-moderators-setup}

df_Low <- df_moderation %>%
  filter(!is.na(Low_percentage))

# Calculate Low_n as (percentage / 100) * sample size
df_Low$Low_n <-
  round((df_Low$Low_percentage / 100) * df_Low$Sample_Size)
```

### Descriptives

We provide additional descriptives based on the Low trajectory variable, using the full dataset.

```{r Low-descriptives}

# Total number of individuals classified as Low (ignoring missingness in other variables)
total_Low_n <- sum(df_Low$Low_n, na.rm = TRUE)

# Number of unique samples
unique_samples <- length(unique(df_Low$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in Low trajectory", "Number of unique samples"),
  Value = format(c(total_Low_n, unique_samples), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for Low Trajectory Analysis")
```

### Continuous moderators

```{r Low-mods-cont}
fit_cont_Low <- \(data, var) fit_cont(data, var, success_col = "Low_n", denom_col = "Sample_Size")

fits_cont_Low <- map_dfr(cont_vars, ~fit_cont_Low(df_Low, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# table rendering
tbl <- fits_cont_Low %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

tbl_utf8 %>%
  gt() %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r Low-mods-cat}
fit_cat_Low <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "Low_n", denom_col = "Sample_Size")

fits_cat_Low <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_Low, c(list(data = df_Low), spec))
)

# table
## set reference level
ref_prev_low <- fits_cat_Low %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

# setup table
Low_cat_table <- fits_cat_Low %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for LOW
  dplyr::left_join(ref_prev_low, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

# Render
Low_cat_table %>%
  gt() %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```

## 🌿 Decreasing Symptom Trajectory

```{r Decreasing-moderators-setup}
df_Decreasing <- df_moderation %>%
  filter(!is.na(Decreasing_percentage))

# Calculate Decreasing_n as (percentage / 100) * sample size
df_Decreasing$Decreasing_n <-
  round((df_Decreasing$Decreasing_percentage / 100) * df_Decreasing$Sample_Size)
```

### Descriptives

We provide additional descriptives based on the Decreasing trajectory variable, using the full dataset.

```{r Decreasing-descriptives}
# Total number of individuals classified as Decreasing (ignoring missingness in other variables)
total_Decreasing_n <- sum(df_Decreasing$Decreasing_n, na.rm = TRUE)

# Number of unique samples
unique_samples <- length(unique(df_Decreasing$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in Decreasing trajectory", "Number of unique samples"),
  Value = format(c(total_Decreasing_n, unique_samples), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for Decreasing Trajectory Analysis")
```

### Continuous moderators

```{r Decreasing-mods-cont}
fit_cont_Decreasing <- \(data, var) fit_cont(data, var, success_col = "Decreasing_n", denom_col = "Sample_Size")

fits_cont_Decreasing <- map_dfr(cont_vars, ~fit_cont_Decreasing(df_Decreasing, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# table 
# build table
tbl <- fits_cont_Decreasing %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

# render
tbl_utf8 %>%
  gt() %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r Decreasing-mods-cat}
fit_cat_Decreasing <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "Decreasing_n", denom_col = "Sample_Size")

fits_cat_Decreasing <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_Decreasing, c(list(data = df_Decreasing), spec))
)

# table
## build table 
ref_prev_Decreasing <- fits_cat_Decreasing %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

Decreasing_cat_table <- fits_cat_Decreasing %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for Decreasing
  dplyr::left_join(ref_prev_Decreasing, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

## Render
Decreasing_cat_table %>%
  gt() %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```

## ⚠️ Increasing Symptom Trajectory

```{r Increasing-moderators-setup}
df_Increasing <- df_moderation %>%
  filter(!is.na(Increasing_percentage))

# Calculate Increasing_n as (percentage / 100) * sample size
df_Increasing$Increasing_n <-
  round((df_Increasing$Increasing_percentage / 100) * df_Increasing$Sample_Size)
```

### Descriptives

We provide additional descriptives based on the Increasing trajectory variable, using the full dataset.

```{r Increasing-descriptives}
# Total number of individuals classified as Increasing (ignoring missingness in other variables)
total_Increasing_n <- sum(df_Increasing$Increasing_n, na.rm = TRUE)

# Number of unique samples
unique_samples <- length(unique(df_Increasing$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in Increasing trajectory", "Number of unique samples"),
  Value = format(c(total_Increasing_n, unique_samples), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for Increasing Trajectory Analysis")
```

### Continuous moderators

```{r Increasing-mods-cont}
fit_cont_Increasing <- \(data, var) fit_cont(data, var, success_col = "Increasing_n", denom_col = "Sample_Size")

fits_cont_Increasing <- map_dfr(cont_vars, ~fit_cont_Increasing(df_Increasing, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# build table
tbl <- fits_cont_Increasing %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

tbl_utf8 %>%
  gt() %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r Increasing-mods-cat}
fit_cat_Increasing <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "Increasing_n", denom_col = "Sample_Size")

fits_cat_Increasing <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_Increasing, c(list(data = df_Increasing), spec))
)

# build table
ref_prev_Increasing <- fits_cat_Increasing %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

Increasing_cat_table <- fits_cat_Increasing %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for Increasing
  dplyr::left_join(ref_prev_Increasing, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

# render
Increasing_cat_table %>%
  gt() %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```

## 🩸 High Symptom Trajectory

```{r High-moderators-setup}
df_High <- df_moderation

df_High <- df_High %>%
  filter(!is.na(High_percentage))

# Calculate High_n as (percentage / 100) * sample size
df_High$High_n <-
  round((df_High$High_percentage / 100) * df_High$Sample_Size)
```

### Descriptives

We provide additional descriptives based on the High trajectory variable, using the full dataset.

```{r High-descriptives}

# Total number of individuals classified as High (ignoring missingness in other variables)
total_High_n <- sum(df_High$High_n, na.rm = TRUE)

# Number of unique samples
unique_samples <- length(unique(df_High$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in High trajectory", "Number of unique samples"),
  Value = format(c(total_High_n, unique_samples), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for High Trajectory Analysis")
```

### Continuous moderators

```{r High-mods-cont}
fit_cont_High <- \(data, var) fit_cont(data, var, success_col = "High_n", denom_col = "Sample_Size")

fits_cont_High <- map_dfr(cont_vars, ~fit_cont_High(df_High, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# build table
tbl <- fits_cont_High %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

# render
tbl_utf8 %>%
  gt() %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r High-mods-cat}
fit_cat_High <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "High_n", denom_col = "Sample_Size")

fits_cat_High <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_High, c(list(data = df_High), spec))
)

# build table
ref_prev_High <- fits_cat_High %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

High_cat_table <- fits_cat_High %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for High
  dplyr::left_join(ref_prev_High, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

# render
High_cat_table %>%
  gt() %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```

## 🌗 Moderate Symptom Trajectory

```{r Moderate-moderators-setup}
df_Moderate <- df_moderation

df_Moderate <- df_Moderate %>%
  filter(!is.na(Moderate_percentage))

# Calculate Moderate_n as (percentage / 100) * sample size
df_Moderate$Moderate_n <-
  round((df_Moderate$Moderate_percentage / 100) * df_Moderate$Sample_Size)
```

### Descriptives

We provide additional descriptives based on the Moderate trajectory variable, using the full dataset.

```{r Moderate-descriptives}
# Total number of individuals classified as Moderate (ignoring missingness in other variables)
total_Moderate_n <- sum(df_Moderate$Moderate_n, na.rm = TRUE)

# Number of unique samples
unique_samples <- length(unique(df_Moderate$Study))

# Display as a table
additional_info <- data.frame(
  Description = c("Total number of individuals in Moderate trajectory", "Number of unique samples"),
  Value = format(c(total_Moderate_n, unique_samples), big.mark = ",")
)

knitr::kable(additional_info, caption = "Additional Descriptives for Moderate Trajectory Analysis")
```

### Continuous moderators

```{r Moderate-mods-cont}
fit_cont_Moderate <- \(data, var) fit_cont(data, var, success_col = "Moderate_n", denom_col = "Sample_Size")

fits_cont_Moderate <- map_dfr(cont_vars, ~fit_cont_Moderate(df_Moderate, .x)) %>%
  mutate(across(where(is.numeric), ~round(.x, 3)))

# build table
tbl <- fits_cont_Moderate %>%
  mutate(
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col),`P-value`) %>%
  arrange(Predictor)

# ensure the table and its column names are UTF-8
tbl_utf8 <- tbl
names(tbl_utf8) <- enc2utf8(names(tbl_utf8))
tbl_utf8 <- dplyr::mutate(tbl_utf8, dplyr::across(where(is.character), enc2utf8))

tbl_utf8 %>%
  gt() %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  tab_options(table.width = pct(100))
```

### Categorical moderators

```{r Moderate-mods-cat}
fit_cat_Moderate <- \(data, var, keep=NULL, ref=NULL)
  fit_cat(data, var, keep, ref, success_col = "Moderate_n", denom_col = "Sample_Size")

fits_cat_Moderate <- purrr::map_dfr(
  cat_vars, \(spec) do.call(fit_cat_Moderate, c(list(data = df_Moderate), spec))
)

# build table
ref_prev_Moderate <- fits_cat_Moderate %>%
  dplyr::filter(level == ref_level) %>%
  dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

Moderate_cat_table <- fits_cat_Moderate %>%
  # drop ref–ref rows
  dplyr::filter(level != ref_level) %>%
  # join the correct ref prevalence for Moderate
  dplyr::left_join(ref_prev_Moderate, by = c("moderator", "ref_level")) %>%
  # pretty names for ref and comparison
  dplyr::left_join(lvl_map, by = c("moderator", "ref_level" = "orig")) %>%
  dplyr::rename(pretty_ref = pretty) %>%
  dplyr::left_join(lvl_map, by = c("moderator", "level" = "orig")) %>%
  dplyr::rename(pretty_cmp = pretty) %>%
  dplyr::mutate(
    Predictor                 = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
    `Samples (k)`             = n,
    `Reference (Name)`        = dplyr::coalesce(pretty_ref, ref_level),
    `Reference (Prevalence)`  = sprintf("%.1f%%", ref_prev),
    `Comparison (Name)`       = dplyr::coalesce(pretty_cmp, level),
    `Comparison (Prevalence)` = sprintf("%.1f%%", prevalence_pct),
    !!delta_col               := round(diff_vs_ref_pp, 2),  # <-- key change
    `p-value`                 = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
  ) %>%
  dplyr::select(
    Predictor, `Samples (k)`, `Reference (Name)`, `Reference (Prevalence)`,
    `Comparison (Name)`, `Comparison (Prevalence)`, dplyr::all_of(delta_col), `p-value`
  ) %>%
  dplyr::arrange(Predictor, `Comparison (Name)`)

# render
Moderate_cat_table %>%
  gt() %>%
  cols_width(Predictor ~ px(110)) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything())
    )
  ) %>%
  tab_options(
    data_row.padding = px(4),
    column_labels.padding = px(6),
    table.font.size = px(13)
  )
```
# Substantive Models
## Setup
```{r substantive-mods-setup-functions}
background_vars <- c(
  "Mean_age", "Developmental_age","Percentage_women", "Percentage_minority", "Percentage_partner", "High_education", "Location", "Location_US"
  )

trauma_vars <- c(
  "Discrete", "Health_First", "Trauma_type", "Military", "Occupational_trauma", "Trauma_exposure"
  )

method_vars <- c(
  "Sample_Size_100", "Grolts", "Entropy", "Scale_moderator", "Diagnostic_DSM", "Trajectory_analysis", "TP_assessments", "N_trajectories",
  "Trauma_TP1", "TP1_TPX", "Trauma_TPX"
  )

.ctrl <- ctrl # Use the pre-defined 'ctrl'

# Check if a variable is categorical using the pre-defined list from 'mods-setup-functions'
is_categorical <- function(v) {
  v %in% sapply(cat_vars, `[[`, "var")
}

# get the p-value column name
.pcol <- function(coef_mat) {
  hits <- intersect(colnames(coef_mat), c("Pr(>|z|)", "Pr(>|t|)", "Pr(>|Chisq|)"))
  if (length(hits)) hits[1] else NA_character_
}

# Find keep/ref settings for a categorical variable from cat_vars list
get_cat_specs <- function(var) {
  specs <- Filter(function(x) x$var == var, cat_vars)
  if (length(specs) == 1) return(specs[[1]]) else return(NULL)
}

# Fit a categorical moderator (using get_cat_specs)
fit_cat <- function(data, var, success_col = "Low_n", denom_col = "Sample_Size") {
  specs <- get_cat_specs(var)
  if (is.null(specs)) return(tibble::tibble())

  d <- data %>%
    dplyr::filter(!is.na(.data[[var]])) %>%
    { if (!is.null(specs$keep)) dplyr::filter(., .data[[var]] %in% specs$keep) else . } %>%
    add_binom_cols(success_col, denom_col) %>%
    dplyr::mutate(
      !!rlang::sym(var) := forcats::as_factor(.data[[var]]),
      !!rlang::sym(var) := if (!is.null(specs$ref) && specs$ref %in% levels(.data[[var]])) {
        forcats::fct_relevel(.data[[var]], specs$ref)
      } else {
        forcats::fct_drop(.data[[var]])
      }
    )

  if (nlevels(d[[var]]) < 2) return(tibble::tibble())

  fml <- stats::as.formula(paste0("cbind(Success, Failure) ~ ", var, " + (1 | Study)"))
  m <- lme4::glmer(fml, data = d, family = binomial, control = .ctrl)

  fe <- lme4::fixef(m)
  b0 <- unname(fe["(Intercept)"])
  lvls <- levels(d[[var]])

  add <- sapply(lvls, function(L) { nm <- paste0(var, L); if (nm %in% names(fe)) fe[[nm]] else 0 })

  s <- summary(m)$coefficients
  pcol <- .pcol(s)
  pvals <- sapply(lvls, function(L) {
    nm <- paste0(var, L)
    if (!is.na(pcol) && nm %in% rownames(s)) unname(s[nm, pcol]) else NA_real_
  })

  p <- plogis(b0 + add)
  p_ref <- p[1]

  tibble::tibble(
    trajectory      = sub("_n$", "", success_col),
    moderator       = var,
    n               = stats::nobs(m),
    ref_level       = lvls[1],
    level           = lvls,
    prevalence_pct  = 100 * p,
    diff_vs_ref_pp  = 100 * (p - p_ref),
    p_value         = pvals
  )
}

# Run a set of variables on a trajectory dataset
run_group <- function(df_traj, vars, success_col = "Low_n", denom_col = "Sample_Size") {
  is_cat <- vapply(vars, is_categorical, logical(1))
  cont_vars_group <- vars[!is_cat]
  cat_vars_group  <- vars[is_cat]

  res_cont <- if (length(cont_vars_group)) purrr::map_dfr(cont_vars_group, ~fit_cont(df_traj, .x, success_col, denom_col)) else tibble::tibble()
  res_cat  <- if (length(cat_vars_group))  purrr::map_dfr(cat_vars_group, ~fit_cat (df_traj, .x, success_col, denom_col)) else tibble::tibble()

  list(cont = res_cont, cat = res_cat)
}


# Unified table column definitions
UNIFIED_COLS <- c(
  "Predictor", "k_shared",
  "Ref_Prevalence_Cont", "Delta_Marginal_Cont", "p_value_shared",
  "Ref_Name_Cat", "Ref_Prevalence_Cat", "Comp_Name_Cat", "Comp_Prevalence_Cat", "Delta_Marginal_Cat"
)

.empty_unified_tbl <- function() {
  tibble::tibble(
    Predictor          = character(),
    k_shared           = integer(),
    Ref_Prevalence_Cont  = character(),
    Delta_Marginal_Cont  = character(),
    p_value_shared     = character(),
    Ref_Name_Cat       = character(),
    Ref_Prevalence_Cat   = character(),
    Comp_Name_Cat      = character(),
    Comp_Prevalence_Cat  = character(),
    Delta_Marginal_Cat   = character()
  )
}

.na_label <- "-"

# make_group_table function
make_group_table <- function(group_res, title, na_label = .na_label) {

  # --- A. Continuous ---
  cont_tbl <- if (!is.null(group_res$cont) && nrow(group_res$cont) > 0) {
    group_res$cont %>%
      dplyr::mutate(
        Predictor         = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
        k_shared          = as.integer(Samples),
        Ref_Prevalence_Cont = sprintf("%.2f%%", prevalence_at_0_pct),
        Delta_Marginal_Cont = paste0(round(marginal_delta_pp_per_unit, 2)),
        p_value_shared      = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
        Ref_Name_Cat        = na_label,
        Ref_Prevalence_Cat  = na_label,
        Comp_Name_Cat     = na_label,
        Comp_Prevalence_Cat = na_label,
        Delta_Marginal_Cat  = na_label
      ) %>%
      dplyr::select(dplyr::all_of(UNIFIED_COLS))
  } else {
    .empty_unified_tbl()
  }

  # --- B. Categorical ---
  cat_tbl <- if (!is.null(group_res$cat) && nrow(group_res$cat) > 0) {
    ref_prev <- group_res$cat %>%
      dplyr::filter(level == ref_level) %>%
      dplyr::select(moderator, ref_level, ref_prev = prevalence_pct)

    group_res$cat %>%
      dplyr::filter(level != ref_level) %>%
      dplyr::left_join(ref_prev, by = c("moderator","ref_level")) %>%
      dplyr::left_join(lvl_map, by = c("moderator","ref_level" = "orig")) %>%
      dplyr::rename(pretty_ref = pretty) %>%
      dplyr::left_join(lvl_map, by = c("moderator","level" = "orig")) %>%
      dplyr::rename(pretty_cmp = pretty) %>%
      dplyr::mutate(
        Predictor         = dplyr::recode(moderator, !!!mod_labels, .default = moderator),
        k_shared          = as.integer(n),
        Ref_Prevalence_Cont = na_label,
        Delta_Marginal_Cont = na_label,
        Ref_Name_Cat      = dplyr::coalesce(pretty_ref, ref_level),
        Ref_Prevalence_Cat  = sprintf("%.1f%%", ref_prev),
        Comp_Name_Cat     = dplyr::coalesce(pretty_cmp, level),
        Comp_Prevalence_Cat = sprintf("%.1f%%", prevalence_pct),
        Delta_Marginal_Cat  = sprintf("%.2f", diff_vs_ref_pp),
        p_value_shared    = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value))))
      ) %>%
      dplyr::select(dplyr::all_of(UNIFIED_COLS))
  } else {
    .empty_unified_tbl()
  }


  # ---- C. Bind + render ----
  out <- dplyr::bind_rows(cont_tbl, cat_tbl)

  gt::gt(out) %>%
    gt::tab_header(title = enc2utf8(title)) %>%
    gt::cols_label(
      Predictor = gt::md("**Predictor**"),
      k_shared = gt::md("k"),
      p_value_shared = gt::md("p-value"),
      Ref_Prevalence_Cont = gt::md("Relative<br>prevalence at 0%"),
      Delta_Marginal_Cont = gt::md("Marginal Δ"), 
      Ref_Name_Cat = gt::md("Reference (Name)"),
      Ref_Prevalence_Cat = gt::md("Reference Prevalence"),
      Comp_Name_Cat = gt::md("Comparison (Name)"),
      Comp_Prevalence_Cat = gt::md("Comparison Prevalence"),
      Delta_Marginal_Cat = gt::md("Marginal Δ")
    ) %>%
    gt::tab_spanner(
      label = gt::md("**Continuous Moderators**"),
      columns = c(Ref_Prevalence_Cont, Delta_Marginal_Cont),
      id = "cont_group"
    ) %>%
    gt::tab_spanner(
      label = gt::md("**Categorical Moderators**"),
      columns = c(Ref_Name_Cat, Ref_Prevalence_Cat, Comp_Name_Cat, Comp_Prevalence_Cat, Delta_Marginal_Cat),
      id = "cat_group"
    ) %>%
    gt::cols_align("left",  columns = Predictor) %>%
    gt::cols_align("center", columns = -Predictor) %>%
    gt::cols_width(Predictor ~ gt::px(180), k_shared ~ gt::px(50), p_value_shared ~ gt::px(70)) %>%
    gt::tab_style(
      style = gt::cell_text(weight = "bold"),
      locations = gt::cells_column_labels(columns = gt::everything())
    ) %>%
    gt::tab_options(
      table.width = gt::pct(100),
      data_row.padding = gt::px(4),
      column_labels.padding = gt::px(6),
      table.font.size = gt::px(13),
      table.font.names = "Times New Roman"
    )
}
```

## 🚀 Low Symptom Trajectory

```{r Low-mods-fitting}
# Run the three groups for this trajectory
bg_res_Low <- run_group(df_Low, success_col = "Low_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_Low <- run_group(df_Low, success_col = "Low_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_Low <- run_group(df_Low, success_col = "Low_n", denom_col = "Sample_Size", vars = method_vars)

all_vars <- c(background_vars, trauma_vars, method_vars)
all_Low_res <- run_group(df_Low, success_col = "Low_n", denom_col = "Sample_Size", vars = all_vars)
```

```{r Low-mods-tables}
make_group_table(bg_res_Low, "Low Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_Low, "Low Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_Low, "Low Symptom Trajectory: Methodological Moderators")
make_group_table(all_Low_res, "Low Symptom Trajectory: All Moderators")
```
## ⚠️ Increasing Symptom Trajectory

```{r Increasing-mods-fitting}
# Run the three groups for this trajectory
bg_res_Increasing <- run_group(df_Increasing, success_col = "Increasing_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_Increasing <- run_group(df_Increasing, success_col = "Increasing_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_Increasing <- run_group(df_Increasing, success_col = "Increasing_n", denom_col = "Sample_Size", vars = method_vars)

all_vars <- c(background_vars, trauma_vars, method_vars)
all_Increasing_res <- run_group(df_Increasing, success_col = "Increasing_n", denom_col = "Sample_Size", vars = all_vars)
```

```{r Increasing-mods-tables}
make_group_table(bg_res_Increasing, "Increasing Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_Increasing, "Increasing Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_Increasing, "Increasing Symptom Trajectory: Methodological Moderators")
make_group_table(all_Increasing_res, "Increasing Symptom Trajectory: All Moderators")
```
## 🌿 Decreasing Symptom Trajectory

```{r Decreasing-mods-fitting}
# Run the three groups for this trajectory
bg_res_Decreasing <- run_group(df_Decreasing, success_col = "Decreasing_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_Decreasing <- run_group(df_Decreasing, success_col = "Decreasing_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_Decreasing <- run_group(df_Decreasing, success_col = "Decreasing_n", denom_col = "Sample_Size", vars = method_vars)

all_vars <- c(background_vars, trauma_vars, method_vars)
all_Decreasing_res <- run_group(df_Decreasing, success_col = "Decreasing_n", denom_col = "Sample_Size", vars = all_vars)
```

```{r Decreasing-mods-tables}
make_group_table(bg_res_Decreasing, "Decreasing Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_Decreasing, "Decreasing Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_Decreasing, "Decreasing Symptom Trajectory: Methodological Moderators")
make_group_table(all_Decreasing_res, "Decreasing Symptom Trajectory: All Moderators")
```
## 🩸 High Symptom Trajectory

```{r High-mods-fitting}
# Run the three groups for this trajectory
bg_res_High <- run_group(df_High, success_col = "High_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_High <- run_group(df_High, success_col = "High_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_High <- run_group(df_High, success_col = "High_n", denom_col = "Sample_Size", vars = method_vars)

all_vars <- c(background_vars, trauma_vars, method_vars)
all_High_res <- run_group(df_High, success_col = "High_n", denom_col = "Sample_Size", vars = all_vars)
```

```{r High-mods-tables}
make_group_table(bg_res_High, "High Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_High, "High Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_High, "High Symptom Trajectory: Methodological Moderators")
make_group_table(all_High_res, "High Symptom Trajectory: All Moderators")
```

## 🌗 Moderate Symptom Trajectory

```{r Moderate-mods-fitting}
# Run the three groups for this trajectory
bg_res_Moderate <- run_group(df_Moderate, success_col = "Moderate_n", denom_col = "Sample_Size", vars = background_vars)
tr_res_Moderate <- run_group(df_Moderate, success_col = "Moderate_n", denom_col = "Sample_Size", vars = trauma_vars)
md_res_Moderate <- run_group(df_Moderate, success_col = "Moderate_n", denom_col = "Sample_Size", vars = method_vars)

all_vars <- c(background_vars, trauma_vars, method_vars)
all_Moderate_res <- run_group(df_Moderate, success_col = "Moderate_n", denom_col = "Sample_Size", vars = all_vars)
```

```{r Moderate-mods-tables}
make_group_table(bg_res_Moderate, "Moderate Symptom Trajectory: Socio-Demographic Moderators")
make_group_table(tr_res_Moderate, "Moderate Symptom Trajectory: Trauma-Related Moderators")
make_group_table(md_res_Moderate, "Moderate Symptom Trajectory: Methodological Moderators")
make_group_table(all_Moderate_res, "Moderate Symptom Trajectory: All Moderators")
```

# Model 2: Multivariate BRMS Models with Missing-Data Imputation

The second modelling stage estimates multivariate Bayesian regression models in which all moderators are entered simultaneously using a binomial likelihood. This approach allows us to assess the unique contribution of each predictor while adjusting for the influence of all other moderators in the model. By modelling all predictors jointly, these analyses move beyond simple associations and provide a more rigorous test of independent effects.

Continuous predictors are standardised to place effects on a comparable scale, and categorical predictors are entered as factors with defined reference categories. Missing values in continuous moderators are handled directly within the Bayesian framework using model-based imputation, allowing all available information to be retained rather than relying on listwise deletion. Prior to model fitting, correlation matrices are used to identify highly collinear predictors, with variables exceeding a correlation of .95 excluded to ensure model stability.

For each trajectory (Low, Decreasing, Increasing, High, and Moderate), three nested models are estimated: a background-only model, a model adding trauma-related moderators, and a full model including methodological variables. This hierarchical structure allows us to evaluate whether additional domains of moderators improve explanatory power beyond basic sample characteristics. Model results are summarised using posterior medians and 95% credible intervals, with convergence diagnostics used to verify estimation quality, and predictive performance compared across models using leave-one-out cross-validation.

## Correlation of Continous Predictors

Since the models incorporate multiple moderators simultaneously, it is important to understand how strongly continuous variables relate to one another. High correlations may signal redundancy or confounding. By computing and visualising the correlation matrix, we obtain an initial sense of the structure among moderators, helping us to anticipate potential collinearity and refine the interpretation of subsequent model coefficients.
Here, we first provide the correlation matrix of continuous variables of the whole dataset.

```{r correlation-table,  results='asis', message=FALSE}
# GT Correlation Table for Continuous Predictors
generate_correlation_gt <- function(df, vars = NULL, trajectory = NULL, show_warnings = TRUE) {
  
  # Get continuous variables from MOD_SPECS
  cont_vars <- MOD_SPECS |> 
    purrr::keep(~ .x$type == "cont") |> 
    purrr::map_chr("var")
  
  # Keep only variables present in df
  cont_vars <- intersect(cont_vars, names(df))
  
  # Compute correlation matrix
  cor_mat <- round(
    cor(df[, cont_vars], use = "pairwise.complete.obs"),
    2
  )
  
  # Convert to long format
  cor_df <- cor_mat %>%
    as.data.frame() %>%
    tibble::rownames_to_column("Var1") %>%
    tidyr::pivot_longer(-Var1, names_to = "Var2", values_to = "Correlation")
  
  # Initialize exclusions
  exclude_vars <- character(0)
  
  # high correlation check
  if (show_warnings) {
    
    # Filter for high correlations and exclude self-pairs + duplicates
    high_cor_df <- cor_df %>%
      dplyr::filter(abs(Correlation) > 0.95 & Var1 != Var2) %>%
      dplyr::rowwise() %>%
      dplyr::filter(Var1 < Var2) %>%
      dplyr::ungroup()
    
    # If any high-correlation pairs exist:
    if (nrow(high_cor_df) > 0) {
      
      # Add pretty labels
      high_cor_df$Pretty1 <- MOD_LABELS[high_cor_df$Var1]
      high_cor_df$Pretty2 <- MOD_LABELS[high_cor_df$Var2]
      
      # Define exclusions for downstream analysis
      exclude_vars <- unique(c(high_cor_df$Var1, high_cor_df$Var2))
      
      # Print warning message
      cat("\n")
      cat("--- WARNING: HIGHLY CORRELATED VARIABLES (Correlation > 0.95) ---\n")
      cat(
        paste0(
          "The following variable pairs have a correlation greater than 0.95 and ",
          "will be candidates for **exclusion** from the analysis for the '", 
          trajectory, "' trajectory:\n"
        )
      )
      
      # Print table
      print(
        high_cor_df %>%
          dplyr::select(Pretty1, Pretty2, Correlation) %>%
          dplyr::rename(Variable_1 = Pretty1, Variable_2 = Pretty2) %>%
          data.frame()
      )
      
      cat("\nVariables flagged for exclusion:\n")
      print(exclude_vars)
      cat("\n")
      
    } else {
      cat("\n--- NOTE: No variable pairs found with correlation > 0.95. ---\n\n")
    }
  }
  
  # Add pretty labels for GT table
  cor_df$Pretty1 <- MOD_LABELS[cor_df$Var1]
  cor_df$Pretty2 <- MOD_LABELS[cor_df$Var2]
  
  # 6. Pivot to wide format
  cor_wide <- cor_df %>%
    dplyr::select(Pretty1, Pretty2, Correlation) %>%
    tidyr::pivot_wider(
      names_from = Pretty2,
      values_from = Correlation
    ) %>%
    dplyr::rename(Variable = Pretty1)
  
  # Build GT table
  cor_gt <- cor_wide %>%
    gt::gt(rowname_col = "Variable") %>%
    gt::tab_header(
      title = paste0(trajectory, " Trajectory Correlation Matrix: Continuous Variables"),
      subtitle = "Pearson correlations (pairwise complete observations)"
    ) %>%
    gt::fmt_number(columns = gt::everything(), decimals = 2) %>%
    gt::data_color(
      columns = gt::everything(),
      colors = scales::col_numeric(
        palette = c("red", "white", "red"),
        domain = c(-1, 1)
      )
    ) %>%
    gt::tab_options(
      table.font.size = 12,
      data_row.padding = gt::px(3),
      column_labels.font.weight = "bold",
      table.font.names = "Times New Roman"
    )
}

correlations_df <- generate_correlation_gt(df_moderation, show_warnings = FALSE)
correlations_df
```

Due to high collinearity, the variable "Time span of the study (months)" (a.k.a. TP1-TPX) will be excluded from all analyses for this Bayesian model. Trajectory-specific correlation matrices are provided in the corresponding sections of this model and variables that have a correlation index of > .95 are excluded from the analyses. 

## Data Preperation for BRMS

Before fitting multivariate models, the data are prepared for Bayesian estimation and missing-data imputation. Continuous moderators are standardised to place them on a comparable scale, and categorical variables are recoded to ensure proper factor structure. This preparation ensures stable estimation and interpretable coefficients in the multivariate analyses.

```{r BRMS-data-prep}
# Function to apply model specifications to a data frame
apply_mod_specs <- function(df, mod_specs) {
  
  # Iterate over each specification
  for (spec in mod_specs) {
    
    var  <- spec$var
    type <- spec$type
    
    # Skip if variable is not in the data frame
    if (!var %in% names(df)) next
    
    
    # Handle continuous variables
    if (identical(type, "cont")) {
      
      x_raw <- df[[var]]
      # Coerce to numeric
      x_num <- suppressWarnings(as.numeric(x_raw))
      
      # Warning check
      if (all(is.na(x_num)) && any(!is.na(x_raw))) {
        warning("Variable '", var, "' became all NA when coerced to numeric.")
      }
      
      df[[var]] <- x_num
      next
    }
    
    
    # Handle categorical variables
    if (identical(type, "cat")) {
      
      # Clean string values
      x <- df[[var]] |>
        as.character() |>
        trimws() |>
        stringr::str_squish() 

      
      # Apply level mapping
      if (!is.null(spec$levels)) {
        lvl_map  <- spec$levels
        raw_keys <- names(lvl_map)
        
        # Apply mapping
        matched  <- x %in% raw_keys
        x[matched] <- lvl_map[x[matched]]
        
        # Report unmatched raw values
        unmatched_raw <- sort(unique(x_raw[!(x_raw %in% raw_keys) & !is.na(x_raw)]))
        if (length(unmatched_raw) > 0L) {
          message(
            "apply_mod_specs: in variable '", var,
            "', some raw values were not mapped by 'levels': ",
            paste(unmatched_raw, collapse = ", ")
          )
        }
      }
      
      
      # Filter to keep only specified levels/labels
      if (!is.null(spec$keep)) {
        if (!is.null(spec$levels)) {
          # Get the mapped labels
          keep_labels <- unname(spec$levels[spec$keep])
        } else {
          # Use 'keep' values directly
          keep_labels <- spec$keep
        }
        
        # Set non-kept values to NA
        x[!(x %in% keep_labels)] <- NA
      }
      
      
      # Convert to factor
      x <- factor(x)
      levels(x) <- trimws(levels(x))
      
      # Set reference level
      if (!is.null(spec$ref)) {
        ref_label <- spec$ref
        
        # Use mapped label for reference if available
        if (!is.null(spec$levels) && spec$ref %in% names(spec$levels)) {
          ref_label <- spec$levels[[spec$ref]]
        }
        
        # Relevel
        if (ref_label %in% levels(x)) {
          x <- stats::relevel(x, ref = ref_label)
        } else {
          warning(
            "Reference level '", spec$ref,
            "' (mapped to '", ref_label,
            "') not found in levels of '", var, "'."
          )
        }
      }
      
      df[[var]] <- x
    }
  }
  
  df
}

# Function to prepare data for brms (apply specs and scale/factor)
prepare_data_brms <- function(df, mod_specs = MOD_SPECS, exclude = NULL) {
  
  # Remove excluded variables from df entirely
  if (!is.null(exclude)) {
    df <- df %>% dplyr::select(-dplyr::all_of(exclude))
  }
  
  # Extract names of continuous variables
  cont_vars <- purrr::map_chr(
    purrr::keep(mod_specs, ~ .x$type == "cont"),
    "var"
  )
  
  # Extract names of categorical variables
  cat_vars <- purrr::map_chr(
    purrr::keep(mod_specs, ~ .x$type == "cat"),
    "var"
  )
  
  # Identify variables present in df
  cont_vars_in_df <- intersect(cont_vars, names(df))
  cat_vars_in_df  <- intersect(cat_vars,  names(df))
  
  # Apply all variable specifications
  df <- apply_mod_specs(df, mod_specs)
  
  # Scale continuous variables
  if (length(cont_vars_in_df) > 0L) {
    df <- df %>%
      dplyr::mutate(
        dplyr::across(
          dplyr::all_of(cont_vars_in_df),
          ~ {
            v <- .
            
            v_num <- suppressWarnings(as.numeric(v))
            
            # Warning check
            if (all(is.na(v_num)) && any(!is.na(v))) {
              warning("After coercion, variable '", dplyr::cur_column(),
                      "' is all NA; scaling will return all NA.")
            }
            
            # Center and scale
            as.numeric(scale(v_num))
          }
        )
      )
  }
  
  # Ensure categorical variables are factors
  if (length(cat_vars_in_df) > 0L) {
    df <- df %>%
      dplyr::mutate(
        dplyr::across(
          dplyr::all_of(cat_vars_in_df),
          ~ factor(.)
        )
      )
  }
  
  # Drop unused levels
  df <- droplevels(df)
  
  df
}
```

## BMRS setup

The Bayesian models are implemented using the brms package, which provides flexible model specification and seamless integration of missing-data imputation via the mi() syntax. The chosen sampling settings (2 chains, 3000 iterations) balance computational feasibility with the need for stable posterior estimates.

```{r BRMS-setup}
brms_args <- list(
  family    = binomial(link = "logit"),
  chains    = 2,
  iter      = 3000,
  warmup    = 1000,
  cores     = 8,
  seed      = 1,
  backend   = "cmdstanr",
  refresh   = 0,
  control   = list(adapt_delta = .85)
)
```

## Portable Functions for Model Fitting

This section defines functions that construct the multivariate models automatically for each trajectory. Importantly, missing predictors are estimated through univariate imputation. This approach avoids the loss of information that would occur with listwise deletion and enables the inclusion of a broad predictor set even when individual studies did not report all moderators.

### Fitting the Multivariate Bayesian Models

The multivariate models move beyond simple associations and estimate the unique contribution of each predictor while adjusting for all others in the model. We estimate three nested models per trajectory:

M1: Background-variables-only model

M2: Background + Trauma-related variables model

M3: Full model (Background + Trauma + Methodological variables)

Comparing these models helps determine whether trauma-related or methodological factors contribute additional explanatory value beyond basic sample characteristics.
These groups are widely the same across trajectories, but may differ slightly based on exclusions due to collinearity.

```{r BRMS-model-fitting-functions}
## Build predictor families
build_predictor_families <- function(df, cont_vars, cat_vars) {
  
  fams <- setNames(
    rep(list(gaussian()), length(cont_vars)),
    cont_vars
  )
  
  for (v in cat_vars) {
    if (!v %in% names(df)) next
    
    x <- factor(df[[v]])
    nlev <- nlevels(x)
    
    if (nlev < 2) {
      stop("Variable ", v, " has <2 levels after cleaning.")
    }
    
    # Use categorical() for ALL categorical predictors (2+ levels)
    fams[[v]] <- categorical()
  }
  
  fams
}


## univariate imputation formula 
fit_brms_models_mi <- function(df,
                               success_col_name,
                               model_vars,
                               cont_vars,
                               brms_args) {
  if (!success_col_name %in% names(df)) {
    stop("Outcome variable '", success_col_name, "' not found in df.")
  }
  
  model_vars <- intersect(model_vars, names(df))
  model_vars <- setdiff(model_vars, success_col_name)
  
  if (length(model_vars) == 0L) {
    stop("No valid predictor variables found in 'model_vars'.")
  }
  
  # vars that have missingness
  vars_with_na_raw <- model_vars[
    sapply(df[, model_vars, drop = FALSE], function(x) any(is.na(x)))
  ]
  
  # continuous vars are MI-eligible in brms
  vars_with_na <- intersect(vars_with_na_raw, cont_vars)
  
  vars_without_na <- setdiff(model_vars, vars_with_na_raw)
  
  rhs_terms <- c(
    if (length(vars_with_na)) paste0("mi(", vars_with_na, ")"),
    vars_without_na
  )
  
  main_formula <- bf(
    as.formula(
      paste0(
        success_col_name,
        " | trials(Sample_Size) ~ ",
        paste(rhs_terms, collapse = " + ")
      )
    ),
    family = binomial(link = "logit")
  )
  
  # Case 1: no MI → simple model
  if (length(vars_with_na) == 0L) {
    default_priors <- get_prior(formula = main_formula, data = df)
    fit <- do.call(
      brm,
      c(
        list(
          formula = main_formula,
          data    = df,
          prior   = default_priors
        ),
        brms_args
      )
    )
    return(fit)
  }
  
  # Case 2: MI for continuous vars
  mi_formulas <- lapply(vars_with_na, function(v) {
    bf(as.formula(paste0(v, " | mi() ~ 1")), family = gaussian())
  })
  
  full_formula <- Reduce(`+`, c(list(main_formula), mi_formulas))
  
  default_priors <- get_prior(formula = full_formula, data = df)
  
  fit <- do.call(
    brm,
    c(
      list(
        formula = full_formula,
        data    = df,
        prior   = default_priors
      ),
      brms_args
    )
  )
  
  fit
}

# looping logic for running all models for one trajectory
fit_brms_trajectory_models <- function(df,
                                       success_col_name,
                                       cont_vars,
                                       cat_vars,
                                       mod_specs,
                                       background_vars,
                                       trauma_vars,
                                       method_vars,
                                       brms_args,
                                       exclude_vars = NULL) {
  
  message("Starting fitting for trajectory with outcome: ", success_col_name)
  
  # multicollinearity filtering step
  
  if (!is.null(exclude_vars) && length(exclude_vars) > 0) {
    message("\nExcluding variables due to high correlations:\n",
            paste(" -", exclude_vars, collapse = "\n"))
  }
  
  cont_vars       <- setdiff(cont_vars, exclude_vars)
  cat_vars        <- setdiff(cat_vars, exclude_vars)
  background_vars <- setdiff(background_vars, exclude_vars)
  trauma_vars     <- setdiff(trauma_vars, exclude_vars)
  method_vars     <- setdiff(method_vars, exclude_vars)

  # Model variable setups
  
  model_list <- list(
    model_1_bg        = background_vars,
    model_2_bg_trauma = c(background_vars, trauma_vars),
    model_3_full      = c(background_vars, trauma_vars, method_vars)
  )

  # Fit the models
  
  purrr::map(
    model_list,
    ~ fit_brms_models_mi(
      df               = df,
      success_col_name = success_col_name,
      model_vars       = .x,
      cont_vars        = cont_vars,
      brms_args        = brms_args
    )
  )
}
```

### Summary Tables

We translate posterior distributions into clear summaries that report the median effect size and 95% credible interval for each predictor. 
Predictors whose credible intervals exclude zero are flagged as potentially important. Presenting all three models side-by-side allows readers to see how effects evolve as more predictors are added.

Model diagnostics summarise convergence metrics such as R-hat and effective sample sizes. These indicators ensure that posterior estimates are reliable and that inferences from the multivariate models can be trusted.

Finally, the multivariate models are compared using leave-one-out cross-validation (LOO-CV). This provides an estimate of the relative predictive accuracy of the three model specifications and helps assess whether additional moderator domains improve the overall model fit for each trajectory. LOO-CV ranks models by how well they predict unseen data using ELPD, where higher ELPD values indicate better predictive accuracy. The best model has ΔELPD = 0, and other models have more negative ΔELPD values the worse they perform, with the SE showing how confidently we can distinguish those differences. Optional model weights summarize each model’s relative predictive support, where higher weights signal a model that is more likely to generalize well to new data.

```{r BRMS-gt-table}
# Predictor Summary Table Function
generate_predictor_table_gt <- function(fitted_models,
                                        success_col_name,
                                        trajectory_name,
                                        mod_specs = MOD_SPECS) {
  
  var_names  <- purrr::map_chr(mod_specs, "var") # Extract variable names from specifications
  var_labels <- purrr::map_chr(mod_specs, "label") |> # Extract variable labels
    stats::setNames(var_names) # Name labels with variable names
  
  find_base_var <- function(t0) {

    # Order candidate variable names from longest to shortest
    hits <- var_names[order(nchar(var_names), decreasing = TRUE)]
  
    # Keep only those whose names appear in the term
    hit <- purrr::keep(hits, ~ grepl(.x, t0))
  
    # If nothing matches, return original term
    if (length(hit) == 0) return(t0)
  
    # Return the first (i.e., longest match)
    hit[[1]]
  }

  all_coefs <- purrr::map_dfr(names(fitted_models), function(model_name) { # Iterate over fitted models
    coefs <- try(brms::fixef(fitted_models[[model_name]], summary = TRUE), silent = TRUE) # Get fixed effects summary
    if (inherits(coefs, "try-error") || length(coefs) == 0) return(NULL) # Skip if error or no coefficients
    
    as.data.frame(coefs) %>%
      tibble::rownames_to_column(var = "Term") %>% # Convert row names (terms) to a column
      dplyr::rename( # Rename columns for clarity
        Median   = Estimate,
        CI_Lower = Q2.5,
        CI_Upper = Q97.5
      ) %>%
      dplyr::mutate(Model = model_name) # Add model name column
  })
  
  coefficient_data <- all_coefs %>%
    dplyr::filter(!grepl("Intercept", Term)) %>% # Remove intercept terms
    dplyr::mutate(
      Term = gsub("^Lown_|^Decreasingn_|^Increasingn_|^Highn_|^Moderaten_", "", Term), # Clean up term names (remove prefixes)
      Term = sub("^mi_", "", Term),
      Term = sub("^mi", "", Term)
 # Remove 'mi' (multiple imputation indicator)
    ) %>%
    dplyr::rowwise() %>% # Operate row by row
    dplyr::mutate(
      BaseVar = find_base_var(Term), # Find the base variable name
      Level   = trimws(sub(paste0("^", BaseVar), "", Term)), # Extract the level/suffix of the variable
      VarLabel = ifelse(BaseVar %in% names(var_labels), # Get the descriptive label for the base variable
                        var_labels[BaseVar],
                        BaseVar),
      Predictor = ifelse( # Create the final predictor label
        Level == "" | is.na(Level),
        VarLabel,
        paste0(VarLabel) # If Level exists, combine it with VarLabel
      )
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      Effect = paste0( # Format effect size and 95% CI
        sprintf("%.2f", Median), " [",
        sprintf("%.2f", CI_Lower), ", ",
        sprintf("%.2f", CI_Upper), "]"
      ),
      Important = ifelse(CI_Lower > 0 | CI_Upper < 0, "Yes", "No"), # Check if CI excludes zero
      Model = dplyr::recode( # Recode model names for display
        Model,
        "model_1_bg"        = "M1 (Background)",
        "model_2_bg_trauma" = "M2 (Background + Trauma)",
        "model_3_full"      = "M3 (Background + Trauma + Methodological)"
      )
    ) %>%
    dplyr::select(Predictor, Model, Effect, Important) # Select final columns
  
  coefficient_wide <- coefficient_data %>%
    tidyr::pivot_wider( # Reshape data from long to wide format
      names_from  = Model, # Columns created based on Model
      values_from = c(Effect, Important), # Values to spread across new columns
      names_glue  = "{.value}_{Model}" # Naming convention for new columns
    )
  
  effect_cols     <- grep("^Effect_", names(coefficient_wide), value = TRUE) # Identify Effect columns
  importance_cols <- grep("^Important_", names(coefficient_wide), value = TRUE) # Identify Important columns
  
  effect_labels <- stats::setNames( # Create labels for Effect columns
    gsub("Effect_", "Effect [95% CI] – ", effect_cols),
    effect_cols
  )
  
  importance_labels <- stats::setNames( # Create labels for Important columns
    gsub("Important_", "Important? – ", importance_cols),
    importance_cols
  )
  
  coefficient_wide %>%
    gt::gt(rowname_col = "Predictor") %>% # Initialize gt table, setting Predictor as row names
    gt::tab_header( # Add table title and subtitle
      title    = gt::md(glue::glue("**Predictor Summary: {trajectory_name} Trajectory**")),
      subtitle = "Median, 95% Credible Interval, and Predictor Importance"
    ) %>%
    gt::cols_label( # Apply custom column labels
      !!!effect_labels,
      !!!importance_labels
    ) %>%
    gt::data_color( # Color cells based on 'Important' status
      columns = dplyr::all_of(importance_cols),
      colors  = function(x) dplyr::case_when(
        x == "Yes" ~ "#d4edda", # Light green for 'Yes'
        x == "No"  ~ "white",   # White for 'No'
        TRUE       ~ "white"
      )
    ) %>%
    gt::tab_options(
      table.font.size = 12,
      data_row.padding = gt::px(3),
      column_labels.font.weight = "bold",
      table.font.names = "Times New Roman"
    )
}

# Diagnostic Table Function
diagnostic_gt_brms <- function(model,
                               model_name = NULL,
                               trajectory_name = NULL,
                               include = c("fixed", "sigma"),
                               title = "Model Diagnostics",
                               subtitle = NULL,
                               mod_labels = MOD_LABELS) {

  s       <- summary(model) # Get the model summary
  parts <- list() # Initialize list to store results
  
  pretty_from_mod <- function(term) {
    t0 <- gsub("`", "", term)
    t0 <- gsub("^mi\\(|\\)$", "", t0)
  
    # 1. Exact match first
    if (t0 %in% names(mod_labels)) {
      return(mod_labels[[t0]])
    }
  
    # 2. Ordered substring matching (longest names first)
    hits <- names(mod_labels)[order(nchar(names(mod_labels)), decreasing = TRUE)]
    hit  <- purrr::keep(hits, ~ grepl(.x, t0))
  
    if (length(hit) >= 1) {
      return(mod_labels[[hit[1]]])
    }
  
    # 3. Underscore-insensitive matching
    t1 <- gsub("_", "", t0)
    mod_vars_no_underscore <- gsub("_", "", names(mod_labels))
    match_index <- match(t1, mod_vars_no_underscore)
  
    if (!is.na(match_index)) {
      original_var_name <- names(mod_labels)[match_index]
      return(mod_labels[[original_var_name]])
    }
  
    # 4. Default fallback
    return(term)
  }


  clean_sigma_name <- function(term) { # Function to clean and label sigma terms
    t <- term
    t <- gsub("`", "", t)
    
    if (t == "sigma") return("Residual SD") # Label for overall residual standard deviation
    
    if (grepl("^sigma_", t)) { # Label for group-specific residual SD
      var_name <- sub("^sigma_", "", t)
      pretty_label <- pretty_from_mod(var_name)
      return(paste0("Residual SD: ", pretty_label))
    }
    
    if (grepl("^sd_", t)) { # Label for random effects standard deviations
      t1 <- sub("^sd_", "", t)
      parts <- strsplit(t1, "__")[[1]]
      if (length(parts) == 2) {
        pretty_param <- pretty_from_mod(parts[2])
        return(paste0("SD(", parts[1], ": ", pretty_param, ")"))
      }
    }
    
    return(term) # Return original term if no cleaning applied
  }
  
  if ("fixed" %in% include) { # Process fixed effects
    fx <- as.data.frame(s$fixed)
    fx$term <- rownames(fx)
    fx <- fx %>%
      dplyr::filter(!grepl("Intercept", term, ignore.case = TRUE)) # Exclude Intercept
    parts$fixed <- fx %>% dplyr::select(term, Estimate, Est.Error, Rhat, Bulk_ESS, Tail_ESS) # Select diagnostic columns
  }
  
  if ("sigma" %in% include && !is.null(s$spec_pars)) { # Process special parameters (sigma/SD)
    sg <- as.data.frame(s$spec_pars)
    sg$term <- rownames(sg)
    
    sg <- sg %>%
      dplyr::filter(!grepl("Intercept", term, ignore.case = TRUE)) %>%  # Exclude Intercept
      dplyr::mutate(term = purrr::map_chr(term, clean_sigma_name))      # Apply cleaning and labeling
    
    parts$sigma <- sg %>% dplyr::select(term, Estimate, Est.Error, Rhat, Bulk_ESS, Tail_ESS) # Select diagnostic columns
  }
  
  tab <- dplyr::bind_rows(parts, .id = "component") # Combine fixed and sigma parts, adding a component ID
  
  tab <- tab %>%
    dplyr::mutate(
      Predictor = ifelse(
        component == "fixed",
        purrr::map_chr(term, pretty_from_mod), # Get pretty labels for fixed effects
        term # Keep cleaned sigma labels
      )
    ) %>%
    dplyr::select(component, Predictor, Estimate, Est.Error, Rhat, Bulk_ESS, Tail_ESS) # Select final columns
  
  gt::gt(tab, rowname_col = "Predictor", groupname_col = "component") %>% # Initialize gt table with grouping
    gt::tab_header(
      title = paste0(title,
                     if (!is.null(model_name)) paste0(" — ", model_name) else "",
                     if (!is.null(trajectory_name)) paste0(" (", trajectory_name, ")") else ""
      ),
      subtitle = subtitle
    ) %>% # Add title and subtitle
    gt::fmt_number( # Format numeric columns
      columns = c(Estimate, Est.Error, Rhat, Bulk_ESS, Tail_ESS),
      decimals = 3
    ) %>%
    gt::cols_label( # Apply column labels
      Estimate  = "Estimate",
      Est.Error = "Std. Error",
      Rhat      = "Rhat",
      Bulk_ESS  = "Bulk ESS",
      Tail_ESS  = "Tail ESS"
    ) %>%
    gt::opt_row_striping() %>% # Add row striping
    gt::tab_style( # Style column headers
      style     = gt::cell_text(weight = "bold"),
      locations = gt::cells_column_labels()
    ) %>%
    gt::tab_options( # Set table options
      table.font.size  = 12,
      table.font.names = "Times New Roman"
    )
}

# Model Comparison Function
loo_gt_table <- function(
  model_list,
  model_names = NULL,
  response = NULL,
  table_title = "Model Comparison (LOO-CV)",
  add_weights = TRUE,              # compute stacking / Pseudo-BMA weights
  weight_method = c("stacking", "pseudobma")
) {
  
  weight_method <- match.arg(weight_method)

  # Model names
  if (is.null(model_names)) {
    model_names <- paste0("Model ", seq_along(model_list))
  }
  stopifnot(length(model_names) == length(model_list))

  # Compute LOO for each model
  loo_list <- lapply(
    model_list,
    function(m) {
      if (!is.null(response)) loo(log_lik(m, resp = response))
      else loo(log_lik(m))
    }
  )

  # Compare LOO models
  loo_comp <- do.call(loo_compare, loo_list)

  # Prepare the df
  df <- loo_comp %>%
    as.data.frame() %>%
    rownames_to_column("ModelIndex") %>%
    mutate(Model = model_names) %>%
    rename(
      # Only columns present in loo_compare:
      ELPD_diff = elpd_diff,
      SE = se_diff
    ) %>%
    mutate(
      ELPD_diff = round(ELPD_diff, 3),
      SE = round(SE, 3)
    ) %>%
    select(Model, ELPD_diff, SE)

  # Add model weights 
  if (add_weights) {
    if (weight_method == "stacking") {
      w <- loo_model_weights(loo_list, method = "stacking")
    } else if (weight_method == "pseudobma") {
      w <- loo_model_weights(loo_list, method = "pseudobma")
    }

    df$Weight <- round(as.numeric(w), 3)
  }

  # Build gt table 
  gt_tbl <- df %>%
    gt() %>%
    tab_header(
      title = md(paste0("**", table_title, "**")),
      subtitle =
        "Higher ELPD indicates better out-of-sample predictive performance (with model averaging weights)"
    ) %>%
    cols_label(
      Model = "Model",
      ELPD_diff = "Δ ELPD (relative to best model)",
      SE = "SE(Δ)",
      Weight = "Model Weight"
    ) %>%
    fmt_number(
      columns = c(ELPD_diff, SE, dplyr::any_of("Weight")),
      decimals = 3
    ) %>%
    # Highlight best model: highest ELPD (i.e., Δ = 0)
    tab_style(
      style = list(
        cell_fill(color = "#f0f0f0"),
        cell_text(weight = "bold")
      ),
      locations = cells_body(rows = ELPD_diff == 0)
    ) %>%
    opt_row_striping() %>%
    opt_table_outline() %>%
    tab_options(
      table.font.size = 14,
      data_row.padding = px(6),
      heading.background.color = "#e8e8ef"
    )

  return(gt_tbl)
}
```

## 🚀 Low Symptom Trajectory

### Data Preparation and Correlation of Continuous Variables

Before fitting the multivariate models, data are prepared for Bayesian estimation through standardising continuous moderators and recoding categorical variables, ensuring stable estimation and interpretable coefficients.
A correlation matrix for the Low trajectory is used to detect highly collinear variables, with those exceeding a .95 correlation being excluded from analyses.

```{r BRMS-correlations-low, results='asis', message=FALSE}
# Prepare the data (cleaning, standardizing, recoding)
df_Low_clean <- prepare_data_brms(df_Low, exclude = "TP1_TPX")

# Determine Collinearity
correlations_Low <- generate_correlation_gt(df = df_Low_clean, trajectory = "Low")
correlations_Low

exclude_vars_Low <- correlations_Low$exclude_vars
```

### Model Fits and Diagnostics

The analysis of the Low Trajectory multivariate models distils posterior distributions by reporting each predictor’s median effect size and 95% credible interval, highlighting predictors whose intervals do not include zero. Displaying the three models together shows how effects change as additional predictors are introduced. Model diagnostics, including R-hat and effective sample sizes, confirm good convergence and lend confidence to the reliability of the posterior estimates and resulting inferences.

```{r BRMS-fitting-low, results='asis', message=FALSE}
# Fit Models for Low Trajectory
models_Low <- fit_brms_trajectory_models(
  df               = df_Low_clean,
  success_col_name = "Low_n",
  cont_vars        = cont_vars,
  cat_vars         = cat_vars,
  mod_specs        = MOD_SPECS,
  background_vars  = background_vars,
  trauma_vars      = trauma_vars,
  method_vars      = method_vars,
  brms_args        = brms_args,
  exclude_vars     = exclude_vars_Low
)

# Produce predictor summary table
table_low <- generate_predictor_table_gt(
  fitted_models = models_Low,
  trajectory_name = "Low"
)
table_low

# Diagnostics for each submodel
diag_low_m1 <- diagnostic_gt_brms(model = models_Low$model_1_bg,
                   model_name = "M1: Background",
                   trajectory_name = "Low")

diag_low_m2 <- diagnostic_gt_brms(model = models_Low$model_2_bg_trauma,
                   model_name = "M2: Background + Trauma",
                   trajectory_name = "Low")

diag_low_m3 <- diagnostic_gt_brms(model = models_Low$model_3_full,
                   model_name = "M3: Background + Trauma + Methodological",
                   trajectory_name = "Low")

diag_low_m1
diag_low_m2
diag_low_m3
```

### Model Comparison

We compare the multivariate models of the Low trajectory using leave-one-out cross-validation (LOO-CV) to evaluate whether additional moderator domains improve predictive accuracy, ranking models by ELPD (with higher values indicating better performance), interpreting ΔELPD and its SE to gauge relative fit, and using optional model weights to summarize each model’s likelihood of generalizing well to new data.

```{r BRMS-model-comparison-low, results='asis', message=FALSE}
# Compare models using LOO-CV
gt_results_Low <- loo_gt_table(
  model_list = list(
    models_Low$model_1_bg,
    models_Low$model_2_bg_trauma,
    models_Low$model_3_full
  ),
  model_names = c( "M1 (Background)", "M2 (Background + Trauma)", "M3 (Background + Trauma + Methodological)"),
  response = "Lown"
)

gt_results_Low
```

# Tables for Paper
## Continuous Moderators Summary Table

```{r cont-mods-summary-table}
# enforce trajectory order
traj_order <- c("Low","Decreasing","Increasing","High","Moderate")

#define trajectory fits
fits_by_traj <- list(
  Low        = fits_cont_Low,
  Decreasing = fits_cont_Decreasing,
  Increasing = fits_cont_Increasing,
  High       = fits_cont_High,
  Moderate   = fits_cont_Moderate
)

# build combined table
tbl_all <- bind_rows(fits_by_traj, .id = "Trajectory") %>%
  mutate(
    Trajectory = factor(Trajectory, levels = traj_order),
    Predictor = dplyr::recode(predictor, !!!mod_labels, .default = predictor),
    Predictor = factor(Predictor, levels = mod_levels),
    `Prevalence (0%)` = sprintf("%.1f%%", prevalence_at_0_pct),
    !!delta_col := marginal_delta_pp_per_unit,
    `P-value` = paste0(fmt_p(p_value), ifelse(is.na(p_value), "", paste0(" ", sig(p_value)))),
    Slope = slope
  ) %>%
  select(Trajectory, Predictor, Samples, `Prevalence (0%)`, Slope, dplyr::all_of(delta_col), `P-value`) %>%
  arrange(Trajectory, Predictor)

# ensure UTF-8
tbl_all_utf8 <- tbl_all
names(tbl_all_utf8) <- enc2utf8(names(tbl_all_utf8))
tbl_all_utf8 <- mutate(tbl_all_utf8, across(where(is.character), enc2utf8))

# render one big table, grouped by Trajectory
tbl_all_utf8 %>%
  gt(groupname_col = "Trajectory") %>%
  cols_width(Predictor ~ px(220)) %>%
  cols_align(align = "center", columns = -Predictor) %>%
  cols_align(align = "left",   columns = Predictor) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_column_labels(columns = everything()),    # bold headers
      cells_row_groups(groups = everything())         # bold group labels
    )
  ) %>%
  tab_options(table.width = pct(100)) %>%
  opt_table_font(font = "Times New Roman") 

# export to word
df <- tbl_all_utf8

ft <- flextable(df)
ft <- width(ft, j = "Predictor", width = 2.2)
ft <- align(ft, j = setdiff(names(df), "Predictor"), align = "center", part = "all")
ft <- align(ft, j = "Predictor", align = "left", part = "all")
ft <- bold(ft, part = "header")
ft <- font(ft, fontname = "Times New Roman", part = "all")
ft <- autofit(ft)

doc <- read_docx()
doc <- body_add_flextable(doc, ft)
doc <- body_end_section_landscape(doc)
print(doc, target = "Continuous_Moderator_Table.docx")
```

## Categorical Moderators Summary Table

```{r cat-mods-summary-table}
# enforce trajectory order
traj_order <- c("Low","Decreasing","Increasing","High","Moderate")

# find and combine all 'fits_cat_*' data frames
fit_names <- ls(pattern = "^fits_cat_")

if (length(fit_names) == 0) {
  knitr::kable(data.frame(Note = "No categorical moderator results found."))
} else {
  all_fits_cat <- mget(fit_names) %>%
    list_rbind(names_to = "traj") %>%
    mutate(traj = str_remove(traj, "fits_cat_"),
         traj = factor(traj, levels = traj_order))

  # Prepare the final long table data for gt
  long_table_data <- all_fits_cat %>%
    # Add reference prevalence to each row for comparison
    left_join(
      y  = filter(., level == ref_level) %>% select(traj, moderator, ref_prev = prevalence_pct),
      by = c("traj", "moderator")
    ) %>%
    filter(level != ref_level) %>%
    # Create and format the final columns
    mutate(
      Predictor               = recode(moderator, !!!mod_labels, .default = moderator),
      `Reference (Name)`      = coalesce(lvl_map$pretty[match(ref_level, lvl_map$orig)], ref_level),
      `Comparison (Name)`     = coalesce(lvl_map$pretty[match(level, lvl_map$orig)], level),
      `Samples (k)`           = n,
      `Reference Prevalence`  = sprintf("%.1f%%", ref_prev),
      `Comparison Prevalence` = sprintf("%.1f%%", prevalence_pct),
      !!delta_col      := round(diff_vs_ref_pp, 2),
      `p-value`               = paste0(fmt_p(p_value), " ", sig(p_value))
    ) %>%
    # Select and order the final columns
    select(
      traj, # This column is kept for grouping
      Predictor,
      `Samples (k)`,
      `Reference (Name)`,
      `Reference Prevalence`,
      `Comparison (Name)`,
      `Comparison Prevalence`,
      !!delta_col,
      `p-value`
    ) %>%
    arrange(traj, Predictor, `Comparison (Name)`)

  # Build the final table, grouped by trajectory
  long_table_data %>%
    gt(groupname_col = "traj") %>%
    # Apply styling
    cols_width(Predictor ~ px(220)) %>%
    cols_align(align = "center", columns = where(is.numeric) | where(is.character)) %>%
    cols_align(align = "left", columns = c(Predictor, `Reference (Name)`, `Comparison (Name)`)) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      locations = list(
      cells_column_labels(columns = everything()),    # bold headers
      cells_row_groups(groups = everything())         # bold group labels
    )
     ) %>%
  tab_options(table.width = pct(100)) %>%
  opt_table_font(font = "Times New Roman") 
}

# export to word
df2 <-  long_table_data

ft <- flextable(df2)
ft <- width(ft, j = "Predictor", width = 2.2)
ft <- align(ft, j = setdiff(names(df2), "Predictor"), align = "center", part = "all")
ft <- align(ft, j = "Predictor", align = "left", part = "all")
ft <- bold(ft, part = "header")
ft <- font(ft, fontname = "Times New Roman", part = "all")
ft <- autofit(ft)

doc <- read_docx()
doc <- body_add_flextable(doc, ft)
doc <- body_end_section_landscape(doc)
print(doc, target = "Categorical_Moderator_Table.docx")
```
---








